%!TEX root = paper.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Preface}


\textit{This paper was originally written in 2018 as a collaboration between colleagues that are or have previously been at the University of Duisburg-Essen and University of Vienna, but it was never published. It was intended as a continuation of the work started in ``The Prospects of Cloud Gaming: Do the Benefits Outweigh the Costs?'', to explore the reasons why someone would choose a specific video game platform or service.
We think, the results of this survey might today still give interesting insights into the world of engagement and context factors of playing video games, especially when one intends to investigate QoE. We are also still engaged in this line of research, having already published new works in this field, and we strife to continue this work in the future.} 

\textit{
Therefore, in 2021 we wanted to preserve this work in the form it was originally prepared in 2018, with updated affiliations, restored links, and this preface. 
Past reviewers of this manuscript mostly criticized the lack of a clear research question and definition of context. The focus on a very narrow set of context factors could also have been extended to a broader range of factors, especially social factors, and subjected to a deeper analysis. On the other hand, reviewers praised the relevant topic and its investigation through a comparatively large online survey.
}

